Sensitivity of He Flames in X-ray Bursts to Nuclear Physics

### 2023-06-05 xrb5/paper.tex
Add suggestions to xrb5 paper

Since the timescale of accretion between each burst is in the order of $10^4$ secs while the rise time is in the order of $\sim 1 - 10$ sec \citep{Parikh_2013}, it is unlikely for the same thermodynamic condition to exist on the entire surface of the neutron star to start burning simultaneously \citep{Shara_1982}. Therefore, nuclear ignition is likely to begin in a localized region, and spread to the rest of the neutron star \citep{Spitkovsky_2002}. This asymmetrical burning leads to modulations in the observed flux during the rotation of neutron stars, which causes asymmetrical surface brightness during XRBs \citep{strohmayer_2009}. This explains burst oscillation behavior, a millisecond period variation of XRB intensity during rise time, and the double-peak light curve for non-photospheric-radius-expansion XRBs discovered and supported by observational data \citep{Altamirano_2010,Chakraborty_2014,Bhattacharyya_2006, Kaaret_2007,Smith_1997}.

@@

This paper investigates the effect of reaction networks on the flame.  We explore four different networks including a standard 13-isotope alpha chain as our reference network (see Sec.\ \ref{Sec:network}), two
different plasma screening implementations (Sec.\ \ref{Sec:screening}), and how the reactions and hydro are coupled
together (Sec.\ \ref{Sec:integration}).

@@

During an XRB, nucleosynthesis involves more than a thousand isotopes \citep{Woosley_2004,Koike_2004}. However, due to computational constraints, most multi-dimensional simulations incorporate fewer than 20 isotopes. Thus, this manuscript aims to understand how the different approximations used in the nuclear physics impact on the lateral thermonuclear flame in XRBs.

@@

In order to quantify how the flame changes over time, we look at the
density-weighted profile of temperature and energy generation rate in the burning, similar to the procedure in \citet{harpole:2021}. 
In order to filter out the burning regions, we selected cells that are in the 99th percentile for temperature or energy generation rate. A density-weighted average is then applied to temperature and energy generation rate using Equation \ref{Eq:weighted_equation}, where $\left<Q\right>_w$ is the weighted quantity, $C_{99}(Q)$ are cells where the quantity, $Q$, is in the 99th percentile, and $\rho(c_i)$ and $Q(c_i)$ are the density and the quantity in $c_i$ cell.

@@

With {\tt subch\_full}, we see that the ${}^{12}$C is much more depleted and there is much more production of heavy $\alpha$ chain isotopes through the $\alpha$-chain, particularly ${}^{28}$Si, at $t = 20$ ms. This phenomenon is consistent with the results of \cite{Weinberg_2006}. The burning path, ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, provides an efficient route for converting ${}^{12}$C to ${}^{16}$O compared to the $\alpha$-capture rate. 
%Consequently, when $T \sim 1.3 \times 10^9$ K, the flow from ${}^{12}$C to ${}^{16}$O can surpass the triple-$\alpha$ process, resulting in a depletion of ${}^{12}$C and an amplification in the nuclear energy generation rates. 
%Consequently, it leads to the abundant production of heavy elements such as ${}^{28}$Si, as shown in Figure \ref{fig:network_subch_nuc}. 
%The early depletion of ${}^{12}$C also causes a shortage of burning fuels during the later burning stage, corresponding to an overall decline in the energy generation rate following the outburst.

@@

We fit the flame front position with a simple function to estimate
the flame speed.  For the {\tt aprox13} and {\tt subch\_full\_mod} networks, we use the same expression
as in \citet{eiden:2020}:
\begin{equation}\label{Eq:quadratic_fit}
    r(t) = \frac{1}{2}a_0 t^2 + v_0 t + r_0
\end{equation}
For the {\tt subch\_full} and {\tt subch\_simple}, a hyperbolic tangent function is added to the fitting function to account for the burst of acceleration at $10 \mbox{ms} \lesssim t \lesssim 25 \mbox{ms}$.
\begin{equation}\label{Eq:tanh_fit}
    r(t) = A\tanh{\left(\frac{t}{B} + C\right)} + \frac{1}{2}a_0 t^2 + v_0 t + r_0
\end{equation}
Both fitting functions are applied for $t > 8$ ms, and the fitted parameters along with their respective errors are shown in Table \ref{Tab:network_fitted_param}. The errors are calculated by taking the square root of the diagonal of the covariance matrix.

@@

Although a general trend of decreasing $\Gamma_e$ is observed as the flame progress, it is noteworthy that there are more fusion processes involving heavier nuclei at later stages of burning. Assuming triple-$\alpha$ process, it can inferred that $\Gamma \sim 0.05$ and $0.025$ for $t = 50$ ms and $t = 100$ ms. Meanwhile, for oxygen burning at $t = 50$ ms and 100 ms, $\Gamma \sim 0.48$ and $0.32$, respectively. These results indicate that helium burning occurs in the weak screening regime, while heavier nuclei burning processes occur in the intermediate screening regime and gradually transition into the weak regime.

@@

This finding is in agreement with the results from \cite{Chugunov_2007}, where it was shown that the screening effect calculated by \cite{Alastuey_Jancovici:1978} is always greater than \cite{Chugunov_2007} when $\Gamma \lesssim 40$ (See Figure 4 in \cite{Chugunov_2007}). 
%As $\Gamma$ increases, the difference between the two methods gradually decreases (see Figure 4 in \cite{Chugunov_2007}).
It should be emphasized that {\tt SCREEN5} uses calculation routines from \cite{Alastuey_Jancovici:1978} to formulate the intermediate screening function along with the calculations from \cite{Graboske_1973}. Therefore, reactions that fall within the intermediate screening regime experience a slightly higher screening effect from {\tt SCREEN5} compared to {\tt CHUGUNOV2007}. Consequently, the overall energy generation rates in the early time period are slightly higher for {\tt aprox13} compared to {\tt aprox13\_chu}. This phenomenon accelerates the increasing rate of temperature and the energy generation rate in the later stages of burning, amplifying the discrepancy between these two models as time progresses.

>>>>

Since the timescale of accretion between each burst is in the order of $10^4$ secs while the rise 
time is in the order of $\sim 1 - 10$ sec \citep{Parikh_2013}, it is unlikely for the same 
thermodynamic condition to exist on the entire surface of the neutron star to start burning 
simultaneously \citep{Shara_1982}. {\color{red}Previous sentence seems to be missing a few words. 
Could maybe add ``and for the entire surface" before ``to start burning"?} Therefore, nuclear 
ignition is likely to begin in a 
localized region, and spread to the rest of the neutron star \citep{Spitkovsky_2002}. This 
asymmetrical burning leads to modulations in the observed flux during the rotation of neutron 
stars, which causes asymmetrical surface brightness during XRBs \citep{strohmayer_2009}. This 
explains burst oscillation behavior, a millisecond period variation of XRB intensity during rise 
time, and the double-peak light curve for non-photospheric-radius-expansion XRBs discovered and 
supported by observational data \citep{Altamirano_2010,Chakraborty_2014,Bhattacharyya_2006, 
Kaaret_2007,Smith_1997}.

@@

This paper investigates the effect of reaction networks on the flame.  We explore four different networks including a standard 13-isotope alpha chain as our reference network (see Sec.\ \ref{Sec:network}), two
different plasma screening implementations (Sec.\ \ref{Sec:screening}), and how the reactions and 
hydrodynamics are coupled
together (Sec.\ \ref{Sec:integration}).

@@

During an XRB, nucleosynthesis involves more than a thousand isotopes 
\citep{Woosley_2004,Koike_2004}. However, due to computational constraints, most multi-dimensional 
simulations incorporate fewer than 20 isotopes. Thus, this manuscript aims to understand how the 
different approximations used in the nuclear physics impact the lateral thermonuclear flame in XRBs.

@@

In order to quantify how the flame changes over time, we look at the
density-weighted profile of temperature and energy generation rate in the burning, similar to the procedure in \citet{harpole:2021}. 
In order to filter out the burning regions, we selected cells that are in the 99th percentile 
{\color{red} or higher? (assuming same method)} for temperature or energy generation rate. A 
density-weighted average is then applied to temperature and energy generation rate using Equation 
\ref{Eq:weighted_equation}, where $\left<Q\right>_w$ is the weighted quantity, $C_{99}(Q)$ are 
cells where the quantity, $Q$, is in the 99th percentile {\color{red} or higher}, and $\rho(c_i)$ 
and $Q(c_i)$ are the 
density and the quantity in $c_i$ cell.

@@

With {\tt subch\_full}, we see that the ${}^{12}$C is much more depleted {\color{red} in the 
burning region} and there is much more production of heavy $\alpha$ chain isotopes through the 
$\alpha$-chain, particularly ${}^{28}$Si, at $t = 20$ ms. This phenomenon is consistent with the 
results of \cite{Weinberg_2006}. The burning path, ${}^{12}\mbox{C}(\mbox{p}, \gamma) 
{}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, provides an efficient route for converting 
${}^{12}$C to ${}^{16}$O compared to the $\alpha$-capture rate. 
%Consequently, when $T \sim 1.3 \times 10^9$ K, the flow from ${}^{12}$C to ${}^{16}$O can surpass the triple-$\alpha$ process, resulting in a depletion of ${}^{12}$C and an amplification in the nuclear energy generation rates. 
%Consequently, it leads to the abundant production of heavy elements such as ${}^{28}$Si, as shown in Figure \ref{fig:network_subch_nuc}. 
%The early depletion of ${}^{12}$C also causes a shortage of burning fuels during the later burning stage, corresponding to an overall decline in the energy generation rate following the outburst.

@@

We fit the flame front position with a simple function to estimate
the flame speed.  For the {\tt aprox13} and {\tt subch\_full\_mod} networks, we use the same expression
as in \citet{harpole:2021}:
\begin{equation}\label{Eq:quadratic_fit}
    r(t) = \frac{1}{2}a_0 t^2 + v_0 t + r_0
\end{equation}
For the {\tt subch\_full} and {\tt subch\_simple}, a hyperbolic tangent function is added to the fitting function to account for the burst of acceleration at $10 \mbox{ms} \lesssim t \lesssim 25 \mbox{ms}$.
\begin{equation}\label{Eq:tanh_fit}
    r(t) = A\tanh{\left(\frac{t}{B} + C\right)} + \frac{1}{2}a_0 t^2 + v_0 t + r_0
\end{equation}
Both fitting functions are applied for $t > 8$ ms, and the fitted parameters along with their 
respective errors are shown in Table \ref{Tab:network_fitted_param}. The errors are calculated by 
taking the square root of the diagonal of the covariance matrix. {\color{red} It might be worth 
mentioning which fitting function/method was used if that isn't stated anywhere.}

@@

Although a general trend of decreasing $\Gamma_e$ is observed as the flame progress, it is 
noteworthy that there are more fusion processes involving heavier nuclei at later stages of 
burning. Assuming triple-$\alpha$ process, it can inferred that {\color{red}the Coulomb coupling 
parameter (just for clarity)} $\Gamma \sim 0.05$ and $0.025$ for $t = 50$ ms and $t = 100$ ms. 
Meanwhile, for oxygen 
burning at $t = 50$ ms and 100 ms, $\Gamma \sim 0.48$ and $0.32$, respectively. These results 
indicate that helium burning occurs in the weak screening regime, while heavier nuclei burning 
processes occur in the intermediate screening regime and gradually transition into the weak regime.

@@

This finding is in agreement with the results from \cite{Chugunov_2007}, where it was shown that 
the screening effect calculated by \cite{Alastuey_Jancovici:1978} is always greater than 
\cite{Chugunov_2007} when $\Gamma \lesssim 40$ (See Figure 4 in \citealt{Chugunov_2007}). 
%As $\Gamma$ increases, the difference between the two methods gradually decreases (see Figure 4 in \cite{Chugunov_2007}).
It should be emphasized that {\tt SCREEN5} uses calculation routines from \cite{Alastuey_Jancovici:1978} to formulate the intermediate screening function along with the calculations from \cite{Graboske_1973}. Therefore, reactions that fall within the intermediate screening regime experience a slightly higher screening effect from {\tt SCREEN5} compared to {\tt CHUGUNOV2007}. Consequently, the overall energy generation rates in the early time period are slightly higher for {\tt aprox13} compared to {\tt aprox13\_chu}. This phenomenon accelerates the increasing rate of temperature and the energy generation rate in the later stages of burning, amplifying the discrepancy between these two models as time progresses.
